\chapter{Introduction}%
\label{chap:intro}

Monte Carlo (MC) methods have been and still are one of the most
important tools for numerical calculations in particle physics. Be it
for validating the well established Standard Model (SM) or for making
predictions about new theories, MC simulations are the crucial
interface of theory and experimental data, making them directly
comparable.
% Furthermore horizontal scaling is almost trivial to implement in MC
% algorithms, making them well adapted to modern parallel computing.
In this thesis, the use of MC methods will be traced through from
simple integration to the simulation of proton-proton scattering.

The ``guinea pig'' here is the quark-antiquark annihilation into two
photons \(\qqgg\), henceforth called the diphoton process. It forms an
important background to the Higgs decay channel
\(H\rightarrow \gamma\gamma\), which was instrumental in its
discovery, and to a dihiggs decay
\(HH\rightarrow b\bar{b}\gamma\gamma\), a process of recent
interest~\cite{aaboud2018:sf} to study the higgs self coupling and
probe the limits of the SM. All the while the process is still pure
QED at leading order and thus calculable by hand within the scope of
this thesis as is being done in \cref{chap:qqgg}. The obtained result
is compared to the total cross section obtained with the
\sherpa~\cite{Gleisberg:2008ta} event generator, used as matrix
element integrator. In \cref{chap:mc} some simple MC methods are
discussed, implemented and their results compared. After studying some
basic MC integration methods, the \vegas\
algorithm~\cite{Lepage:19781an} is implemented and
evaluated. Subsequently MC sampling methods, which are closely related
to the integration methods, are explored and the output of \vegas\ is
used to improve the sampling efficiency. Histograms of observables are
generated and compared to histograms from \sherpa\ using the
\rivet~\cite{Bierlich:2019rhm} analysis framework. \Cref{chap:pdf}
deals with proton-proton scattering in the partonic picture using
parton density functions, ending with the implementation of a simple
event generator for \(\ppgg\) scattering at \lhc\ conditions. Some
integration and sampling algorithms and their implementation are
adapted to the multidimensional case and histograms of observables are
generated with good efficiency. Because a real \(pp\) scattering event
also entails processes like parton showers, hadronization and
multiple interactions, a realistic simulation should account for those
effects. The impact of those effects on observables is studied in
\cref{chap:pheno} using the \sherpa\ event generator.

\section{Conventions}%
\label{sec:convent}

Throughout natural units with
\(c=1, \hbar = 1, k_B=1, \varepsilon_0 = 1\) are used unless stated
otherwise. The fine structure constant's value \(\alpha = 1/137.036\)
is configured in \sherpa\ and used in analytic calculations.

\section{Source Code}%
\label{sec:source}

The (literate) python code, used to generate most of the results and
figures can be found under
\url{https://github.com/vale981/bachelor_thesis/} and more
specifically in the subdirectory \texttt{prog/python/qqgg}.

The file \texttt{monte\_carlo.py} implements all the monte-carlo
algorithm related functionality as a module. The file
\texttt{analytical\_xs.org} contains a literate computation notebook
that generates all the results of \cref{chap:mc}. The file
\texttt{parton\_density\_function\_stuff.org} contains all the
computations for \cref{chap:pdf}. The python code makes heavy use of
\href{https://www.scipy.org/}{scipy}~\cite{2020Virtanen:Sc} (and of
course \href{https://numpy.org/}{numpy}).

\section{Sherpa Versions}%
\label{sec:versions}

In \cref{sec:compsher,chap:mc} the development version of \sherpa\ has
been used. \Cref{chap:pdf,chap:pheno} use version \texttt{2.2.10} for
reasons of stability.


%%% Local Variables: ***
%%% mode: latex ***
%%% TeX-master: "../document.tex"  ***
%%% End: ***
